Data Analysis and Modeling

### 2018-05-30 scripts/3_data_prep.Rmd
Move revenue estimates to data prep

For the short-term data, we first calculate the total number of bookings, the occupancy rate, blocked rate and proportion of the year over which the property was listed.  Using this data we sum up the total revenue from bookings. For properties that were not listed on the site for the entire year, we impute their annual revenue by dividing the actual revenue by the proportion of the year in which they were listed.^[Properties listed in summer, but not winter may be over imputed and vice versa, however, we expect the biases to cancel out over the entire sample.] Properties with no reserved bookings, and, therefore, no revenue, are removed.

@@

```{r revdensity.plot, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE, fig.width=6,fig.height=3.5, fig.cap='Comparison of Revenue Distributions'}

   # Compute bookings count and str revenue
   countBookings <- function(x){
     length(which(x$status == 'R'))
   } 
   makePotBookings <- function(x){
     round(nrow(x) * (length(which(x$status == 'R')) / 
                        length(which(x$status != 'B'))), 0)
   }

   str_tdf <- str_tdf %>%
     mutate(bookings = unlist(map(daily_data, ~countBookings(.))),
            pot_bookings = unlist(map(daily_data, ~makePotBookings(.))),
            gross_revenue = bookings * med_rate,
            pot_revenue = pot_bookings * med_rate)

   # compute long term revenue
   ltr_tdf <- ltr_tdf %>%
     dplyr::mutate(trans_data = map(trans_data, 
                                    ~mutate(., 
                                            gross_revenue = price * (52 - dom / 7))))

>>>>

```{r ltr_rev}

   ltr_tdf <- ltr_tdf %>%
     dplyr::mutate(trans_data = map(trans_data, 
                                    ~mutate(., 
                                            gross_revenue = price * (52 - dom / 7)))) %>%
     dplyr::mutate(trans_data = map(trans_data, 
                                    ~filter(., gross_revenue >= 5000)))

For the short-term data, we first calculate the total number of bookings, the occupancy rate, blocked rate and proportion of the year over which the property was listed.  Using this data we sum up the total revenue from bookings.

 For properties that were not listed on the site for the entire year, we impute their annual revenue by dividing the actual revenue by the proportion of the year in which they were listed.^[Properties listed in summer, but not winter may be over imputed and vice versa, however, we expect the biases to cancel out over the entire sample.] Properties with no reserved bookings, and, therefore, no revenue, are removed.

@@

```{r obs_bookings, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE}

   # Compute bookings count and str revenue
   countBookings <- function(x){
     length(which(x$status == 'R'))
   }

   str_tdf <- str_tdf %>%
     mutate(bookings = unlist(map(daily_data, ~countBookings(.))),
            occ_rate = bookings / 366,
            revenue = bookings * med_rate)

Next we impute the likely bookings for properties which were not listed on the portal for the entire year.

```{r lik_bookings, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE}

  # Create daily rate summary
  rate_summ <- str_tdf %>% 
   dplyr::select(property_id, daily_data) %>%
    tidyr::unnest() %>%
    dplyr::filter(status != 'B') %>%
    dplyr::group_by(date) %>%
    dplyr::summarize(rate = length(which(status == 'R')) / n(),
                     obs = n())

  # Create new cluster
  abb_cl <- get_default_cluster()

  # Register function and variable
  cluster_copy(abb_cl, impLikelyBookings)
  cluster_copy(abb_cl, rate_summ)
  cluster_library(abb_cl, c("magrittr", 'dplyr', 'tidyr'))

  # Summarize daily by property
  lik_book_df <- str_tdf %>%
    dplyr::filter(total_days != 366) %>%
    multidplyr::partition(property_id, cluster = abb_cl) %>% 
    dplyr::do(impLikelyBookings(., rate_summ = rate_summ)) %>% 
    dplyr::collect()

  # Add to the existing data
  str_tdf <- str_tdf %>%
    left_join(lik_book_df, by='property_id') %>%
    dplyr::mutate(lik_bookings = if_else(is.na(lik_bookings), 
                                         bookings, 
                                         as.integer(lik_bookings)),
                  lik_occ_rate = lik_bookings / 366)

Finally, we compute the potential bookings, assuming that blocked periods were rented at the same rate as the non-block, controlling for dates and property specific performance on non-blocked dates.

```{r pot_bookings, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE}

    # Create new cluster
  abb_cl <- get_default_cluster()

  # Register function and variable
  cluster_copy(abb_cl, impPotentialBookings)
  cluster_copy(abb_cl, rate_summ)
  cluster_library(abb_cl, c("magrittr", 'dplyr', 'tidyr'))

  # Summarize daily by property
  pot_book_df <- str_tdf %>%
    dplyr::filter(block_rate > 0) %>%
    multidplyr::partition(property_id, cluster = abb_cl) %>% 
    dplyr::do(impPotentialBookings(., rate_summ = rate_summ)) %>% 
    dplyr::collect()

  # Add to the existing data
  str_tdf <- str_tdf %>%
    left_join(pot_book_df, by='property_id') %>%
    dplyr::mutate(pot_bookings = if_else(is.na(pot_bookings), 
                                         bookings, 
                                         as.integer(pot_bookings)),
                  pot_bookings = if_else(pot_bookings > 366L, 366L, pot_bookings),
                  pot_occ_rate = pot_bookings / 366)

### Calculating Net Revenues

We begin the analysis by calculating the net revenues for the short term data.  We calculate three types of net revenues, Actual, Likely (Extrapolated) and Potential. The differences are described more fully in the writeup.

Before we start, we estimate two figures necessary to determine utility costs: 1) Costs per person per day (cpppd) and the average number of guests per bedroom. The State of Victoria gives energy costs at \$330 (apartment) to \$520 (house).  For 2 persons in an apartment that works out to \$5 per person per day.  For a house of 4 persons that works out to about \$4 per person per day. We use a mid point value of \$4.50 per person per day.  We use an estimate of 1.5 persons per bedroom in each unit.

```{r utility_estimates, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  # Set cost per person per day
  cpppd <- 4.5

  # Calculate avg number of guests
  str_tdf$avg_guests <- str_tdf$bedrooms * 1.5

We then calculate the actual net revenue by multiplying booking times median rate - utility costs.

```{r act_revenues, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  str_tdf$revenue_net <- ((str_tdf$bookings * str_tdf$med_rate) -
                             (str_tdf$bookings * str_tdf$avg_guests * cpppd))

Next we create Likely Revenue, by imputing occupancy rates and revenues for the remainder of the year for those properties,

```{r lik_revenues, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  # Calculate likely revenue
  str_tdf$lik_revenue_net <- ((str_tdf$lik_bookings * str_tdf$med_rate) -
                                (str_tdf$lik_bookings * str_tdf$avg_guests * cpppd))

Then we create Potential Revenue, by imputing occupancy rates and revenues for blocked periods.

```{r pot_revenues, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  str_tdf$pot_revenue_net <- ((str_tdf$pot_bookings * str_tdf$med_rate) -
                                (str_tdf$pot_bookings * str_tdf$avg_guests * cpppd))

### 2018-05-30 scripts/4_data_analysis.Rmd
Move revenue estimates to data prep

@@

### Calculating Net Revenues

We begin the analysis by calculating the net revenues for the short term data.  We calculate three types of net revenues, Actual, Likely (Extrapolated) and Potential. The differences are described more fully in the writeup.

Before we start, we estimate two figures necessary to determine utility costs: 1) Costs per person per day (cpppd) and the average number of guests per bedroom. The State of Victoria gives energy costs at \$330 (apartment) to \$520 (house).  For 2 persons in an apartment that works out to \$5 per person per day.  For a house of 4 persons that works out to about \$4 per person per day. We use a mid point value of \$4.50 per person per day.  We use an estimate of 1.5 persons per bedroom in each unit.

```{r utility_estimates, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  # Set cost per person per day
  cpppd <- 4.5

  # Calculate avg number of guests
  str_tdf$avg_guests <- str_tdf$bedrooms * 1.5

We then calculate the actual net revenue by multiplying booking times median rate - utility costs.

```{r act_revenues, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  str_tdf$act_revenue <- ((str_tdf$bookings * str_tdf$med_rate) -
                                 (str_tdf$bookings * str_tdf$avg_guests * cpppd))

Next we create Likely Revenue, by imputing occupancy rates and revenues for the remainder of the year for those properties

```{r lik_revenues, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  # Create daily rate summary
  rate_summ <- str_tdf %>% 
   dplyr::select(property_id, daily_data) %>%
    tidyr::unnest() %>%
    dplyr::filter(status != 'B') %>%
    dplyr::group_by(date) %>%
    summarize(rate = length(which(status == 'R')) / n(),
            obs = n())

  # Create new cluster
  abb_cl <- get_default_cluster()

  # Register function and variable
  cluster_copy(abb_cl, impLikelyBookings)
  cluster_copy(abb_cl, rate_summ)
  cluster_library(abb_cl, c("magrittr", 'dplyr', 'tidyr'))

  # Summarize daily by property
  lik_book_df <- str_tdf %>%
    dplyr::filter(total_days != 366) %>%
    multidplyr::partition(property_id, cluster=abb_cl) %>% 
    dplyr::do(impLikelyBookings(., std=g)) %>% 
    dplyr::collect()

  # Add to the existing data
  str_tdf <- str_tdf %>%
    left_join(lik_book_df, by='property_id') %>%
    dplyr::mutate(lik_bookings = if_else(is.na(lik_bookings), bookings, as.integer(lik_bookings)))

  # Calculate likely revenue
  str_tdf$lik_revenue <- ((str_tdf$lik_bookings * str_tdf$med_rate) -
                           (str_tdf$lik_bookings * str_tdf$avg_guests * cpppd))

  # Create Extrapolation parameter
  str_tdf$extr_par <- 366 / str_tdf$total_days

```{r pot_revenues, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  str_tdf$occ_rate <- 1 - (str_tdf$block_rate + str_tdf$avail_rate)

  str_tdf$pot_occ_rate <- ((str_tdf$lik_bookings) / (366 - (366 * str_tdf$block_rate)))

  str_tdf$pot_revenue <- ((str_tdf$pot_occ_rate * 366 * str_tdf$med_rate) -
                           (str_tdf$pot_occ_rate * 366 * str_tdf$avg_guests * cpppd))

@@

  # Extract out DFs (only those that imputed)
  str_tdf <- str_tdf %>%
    dplyr::left_join(imp_df, by='property_id') %>%
    dplyr::mutate(act_pref = ifelse(act_revenue > ltr_imp_revenue, 1, 0),
                  lik_pref = ifelse(lik_revenue > ltr_imp_revenue, 1, 0),
                  pot_pref = ifelse(pot_revenue > ltr_imp_revenue, 1, 0))

@@

  ggplot(str_tdf, aes(x=block_rate, y=nbr_block * extr_par)) + 
    geom_point(fill=str_col, color=str_col, size=.2) +
    ylab('Nbr of Blocked Periods\n') + 
    xlab('\n% of Time Property is Blocked') +
    scale_x_continuous(breaks=seq(0, 1, .2),
                       labels=c('0%', '20%', '40%', '60%', '80%', '100%')) +
    ggtitle('Block Rate vs. Nbr. of Blocked Periods\n') +
    theme(plot.title = element_text(hjust = 0.5))

@@

  str_tdf$blockpertime <- round(str_tdf$nbr_block * str_tdf$extr_par, 1)

@@

  save(str_tdf, ltr_tdf, ltr_df, col_df, suburb_sf, studyarea_sf, sa1_sf, str_sf, ltr_sf,
       pot_glms, pot_ht_glms, lik_glms, lik_ht_glms, core_suburbs,
       file=file.path(data_path, 'analyzed', 'analyzed_data.RData'))

>>>>

  str_col = col_df$rgb[1]

@@

@@

  # Extract out DFs (only those that imputed)
  str_tdf <- str_tdf %>%
    dplyr::left_join(imp_df, by='property_id') %>%
    dplyr::mutate(act_pref = ifelse(revenue_net > ltr_imp_revenue, 1, 0),
                  lik_pref = ifelse(lik_revenue_net > ltr_imp_revenue, 1, 0),
                  pot_pref = ifelse(pot_revenue_net > ltr_imp_revenue, 1, 0))

@@

  ggplot(str_tdf, aes(x=block_rate, y=nbr_block * (366/total_days))) + 
    geom_point(fill=str_col, color=str_col, size=.2) +
    ylab('Nbr of Blocked Periods\n') + 
    xlab('\n% of Time Property is Blocked') +
    scale_x_continuous(breaks=seq(0, 1, .2),
                       labels=c('0%', '20%', '40%', '60%', '80%', '100%')) +
    ggtitle('Block Rate vs. Nbr. of Blocked Periods\n') +
    theme(plot.title = element_text(hjust = 0.5))

@@

  str_tdf$blockpertime <- round(str_tdf$nbr_block * (366 / str_tdf$total_days), 1)

@@

  dir.create(file.path(getwd(), 'data', 'analyzed'))
  save(str_tdf, ltr_tdf, ltr_df, col_df, suburb_sf, studyarea_sf, sa1_sf, str_sf, ltr_sf,
       pot_glms, pot_ht_glms, lik_glms, lik_ht_glms, core_suburbs,
       file=file.path(getwd(), 'data', 'analyzed', 'analyzed_data.RData'))
